surfing

### 2021-06-29 inst/surfing.Rmd
compatriot work

options(width=100)
surfing_table
wildcards <- rownames(surfing_table)[-(1:36)]
wildcards
QS <- rownames(surfing_table)[c(23:33,35:36)]
QS
```

@@

```{r profsuppcalcsurfing2, cache=TRUE}
p <- seq(from=0.0005,to=0.01,by=0.0005)
l <- profsupp(surfing2,"Brazilian_beachtype",p,relative=FALSE)
l0 <- loglik(indep(surfing_maxp),surfing)
```

@@

>>>>

options(width=120)
surfing_table
wildcards <- rownames(surfing_table)[-(1:36)]
wildcards
QS <- rownames(surfing_table)[c(23:33,35:36)]
QS
```

@@

```{r profsuppcalcsurfing2, cache=TRUE}
p <- seq(from=0.001,to=0.01,by=0.0005)
l <- profsupp(surfing2,"Brazilian_beachtype",p,relative=FALSE)
l0 <- loglik(indep(surfing_maxp),surfing)
```

@@

# compatriot selection

```{r calculate_compatriot_monster_LF,cache=FALSE}
scomp <- hyper2()
scomp3 <- hyper2()  # scomp3 is the likelihood function for just the races with 3 competitors, 2 of one nationality and one odd-man-out.
for(v in o){
   r <- v[-1]  # first entry is venue; "r" for "race"
   numberretained <- sum(!(r %in% remove))
   if(numberretained > 1){
     r <- r[!(r %in% remove)]
     if(numberretained ==3){
       nationalities <- c(nationality(r[1]),nationality(r[2]),nationality(r[3]))
       if(length(unique(nationalities)) == 2){
         if(sum(nationalities %in% nationality(r[1])) == 1){ # winner is the odd one out
           Hjj <- pwa(race(r),pwa=r[1])
           print("odd one out is number 1")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[2]) ) == 1){ # second best is the odd one out
           Hjj <- pwa(race(r),pwa=r[2])
           print("odd one out is number 2")
           print(100*surfing_maxp[r])
         } else if(sum(nationalities %in% nationality(r[3])) == 1){  # third best is the odd one out
	   Hjj <- pwa(race(r),pwa=r[3])
           print("odd one out is number 3")
           print(100*surfing_maxp[r])
         } else {
           stop("this cannot happen")
         }
         print(Hjj)
         print("---")
         scomp <- scomp + Hjj      # add compatriot-monster modified likelihood 
         scomp3 <- scomp3 + Hjj
       } else {                    # 3 in race, but no compatriot monster...
         scomp <- scomp + race(r)  # ... so use regular Plackett-Luce for 3-competitor race
       }
     } else {                    # if(numberretained==3) closes...
       scomp <- scomp + race(r)  # numberretained !=3; use regular Plackett-Luce
     }
  } # if(numberretained>1) closes
} # for(v in o) loop closes
```

```{r maximize_compatriot_monster,cache=FALSE}
maxs <- maxp(scomp)
maxs
```

```{r}
pie(maxs)
```

```{r testcompatriotmonster,cache=TRUE}
specificp.gt.test(scomp,"S",0,n=100,maxtry=100)
```

Now profile likelihood:

```{r profsuppcalcscompatriotmonster, cache=TRUE}
p <- c(0.0002,seq(from=0.001,to=0.02,len=10))
l <- profsupp(scomp,"S",p,relative=FALSE)
```

plot it:

```{r}  
plot(p,l,type='b',xlab="compatriot support",ylab="support")
```

```{r}
scomp3
maxp(scomp3)
specificp.gt.test(scomp3,'S',0)
````

We can do a much more basic test on just the entries with three competitors comprising
two of one nationality and one of a different nationality:

```{r basicanalysis}
##     Dora Ferreira   Slater 
##   2.5741  12.5929   3.8933  (3)

##     Smith    Buchan Freestone 
##    5.2764    2.0552    1.8673  (1)

##   Slater   Andino Callinan 
##   3.8933   5.9190   1.8408  (3)

## Crisanto   Wilson   Ibelli 
##   2.1644   2.7100   2.2479  (2)

##     Moniz    Wright Freestone 
##    3.0250    3.6502    1.8673 (1)

## Carmichael  Freestone     Flores 
##     1.8855     1.8673     2.0480 (3)

##    Bourez    Wright Freestone 
##    1.9068    3.6502    1.8673 (1)

## Colapinto    Flores    Bourez 
##    2.9871    2.0480    1.9068  (1)

## Ferreira   Morais   Ibelli 
##  12.5929   1.0817   2.2479 (2)

## Freestone    Flores    Bourez 
##    1.8673    2.0480    1.9068  (1)

##     Dora Ferreira   Morais 
##   2.5741  12.5929   1.0817 (1)

M <- matrix(c(
  2.5741,12.5929,3.8933,3
, 5.2764, 2.0552,1.8673,1
, 3.8933, 5.9190,1.8408,3
, 2.1644, 2.7100,2.2479,2
, 3.0250, 3.6502,1.8673,1
, 1.8855, 1.8673,2.0480,3
, 1.9068, 3.6502,1.8673,1
, 2.9871, 2.0480,1.9068,1
,12.5929, 1.0817,2.2479,2
, 1.8673, 2.0480,1.9068,1
, 2.5741,12.5929,1.0817,1),byrow=TRUE, ncol=4)

for(i in seq_len(nrow(M))){ M[i,1:3] <-   M[i,1:3]/sum(M[i,1:3])}
print(M)  # rows 1-3 of M are strengths of the first, second, third
          # competitors.  Row 4 is the place of the odd-man-out.  Thus
          # on row 1, the odd-man-out came third, and on row 2, the
          # odd-man-out came first.

p <- M[cbind(1:11, M[,4])]
print(p) 
                                        ## element i of p is the probability of the odd-man-out
                                        ## winning, given the maximum likelihood estimate.

ww <-table(replicate(1e5,sum(rbinom(11,1,p))))
print(ww)
sum(ww[7:10])/sum(ww) # pvalue for observation of six wins by the odd-man-out
```
